# Preprocessing

## Load packages

Explicitly load the packages that we need for this analysis.

```{r packages}
library(rio)
library(ck37r)
library(caret)
```

## Load the data

Load the heart disease and iris datasets. 

```{r load_data}
# Load the heart disease dataset using import() from the rio package.
data = import("data-raw/heart.csv")
str(data)
```

## Read background information and variable descriptions  
https://archive.ics.uci.edu/ml/datasets/heart+Disease

```{r}
# Load the iris dataset (built into R)
data(iris)
?iris
str(iris)
```

## Data preprocessing

Data peprocessing is an integral first step in machine learning workflows. Because different algorithms sometimes require the moving parts to be coded in slightly different ways, always make sure you research the algorithm you want to implement so that you properly setup your $y$ and $x$ variables and split your data appropriately. 

One additional preprocessing aspect to consider: datasets that contain factor (categorical) features should typically be expanded out into numeric indicators (this is often referred to as [one-hot encoding](https://hackernoon.com/what-is-one-hot-encoding-why-and-when-do-you-have-to-use-it-e3c6186d008f). You can do this manually with the `model.matrix` R function. This makes it easier to code a variety of algorithms to a dataset as many algorithms handle factors poorly (decision trees being the main exception). Doing this manually is always good practice. In general however, functions like `lm` will do this for you automatically. More on this below. 

> NOTE: Keep in mind that training/test dataset splitting is common, but not always preferred. We will introduce you to cross-validation in the second half of this workshop where _all_ the data are used and multiple training/testing splits are utilized. 

## Handling missing data

Missing values need to be handled somehow. Listwise deletion (deleting any row with at least one missing value) is common but this method throws out a lot of useful information. Many advocate for mean imputation, but arithmetic means are sensitive to outliers. Still, others advocate for Chained Equation/Bayesian/Expectation Maximization imputation (e.g., the [mice](https://www.jstatsoft.org/article/view/v045i03/v45i03.pdf) and [Amelia II](https://gking.harvard.edu/amelia) R packages). K-nearest neighbor imputation can also be useful but median imputation is demonstrated below. However, you will want to learn about [Generalized Low Rank Models](https://stanford.edu/~boyd/papers/pdf/glrm.pdf) for missing data imputation in your research.

First, count the number of missing values across variables in our dataset.  

```{r review_missingness}
colSums(is.na(data))
```

We have no missing values, so let's introduce a few to the "oldpeak" feature for this example to see how it works: 

```{r}
data$oldpeak[c(50, 100, 150, 200, 250)] = NA

# Five missing values in oldpeak in row numbers 50, 100, 150, 200, 250
colSums(is.na(data))
colMeans(is.na(data))
```

There are now 5 missing values in the "oldpeak" feature. Now, median impute the missing values! We also want to create missingness indicators to inform us about the location of missing data. These are additional columns we will add to our data frame that represent the _locations_ within each feature that have missing values - 0 means data are present, 1 means there was a missing (and subsequently imputed) value.  

```{r impute_missing_values}
result = ck37r::impute_missing_values(data, verbose = TRUE, type = "standard")
names(result)

# Use the imputed dataframe.
data = result$data

# View new columns. Note that the indicator feature "miss_oldpeak" has been added as the last column of our data frame. 
str(data)

# No more missing data!
colSums(is.na(data))
```

Since the "ca", "cp", "slope", and "thal" features are currently interger type, convert them to factors. The other relevant variables are either continuous or are already indicators (just 1's and 0's). 

```{r}
data = ck37r::categoricals_to_factors(data,
              categoricals = c("ca", "cp", "slope", "thal"),
              verbose = TRUE)

# Inspect the updated data frame
str(data)
```

## Defining *y* outcome vectors and *x* feature dataframes

## Convert factors to indicators

Now expand "ca", "cp", "slope", and "thal" features to out into indicators. 

```{r factors_to_indicators}
result = ck37r::factors_to_indicators(data, verbose = TRUE)

data = result$data

str(data)
```

## Regression setup

Now that the data have been imputed and properly converted, we can assign the regression outcome variable (`age`) to its own vector for the lasso **REGRESSION tasks**. Remember that lasso can also perform classification as well. 

### Set seed for reproducibility

Take the simple approach to data splitting and divide our data into training and test sets; 70% of the data will be assigned to the training set and the remaining 30% will be assigned to the holdout, or test, set.

### Random versus stratified random split

Since age is a continuous variable and will be the outcome for the OLS and lasso regressions, we will not perform a stratified random split like we will for the classification tasks (see below). Instead, [let's randomly assign](https://stackoverflow.com/questions/17200114/how-to-split-data-into-training-testing-sets-using-sample-function) 70% of the `age` values to the training set and the remaining 30% to the test set.

```{r}
# Create a list to organize our machine learning task.
task_reg = list(
  data = data,
  outcome = "age"
)

# All variables can be used as covariates except the outcome.
(task_reg$covariates = setdiff(names(data), task_reg$outcome))

names(task_reg)

# Define the sizes of training (70%) and test (30%) sets.
(training_size = floor(0.70 * nrow(task_reg$data)))

# Set seed for reproducibility.
set.seed(1)

# Partition the rows to be included in the training set.
training_rows = sample(nrow(task_reg$data), size = training_size)

task_reg$train_rows = training_rows
head(task_reg$train_rows)
```

## Classification setup

Assign the outcome variable to its own vector for the decision tree, random forest, gradient boosted tree, and SuperLearner **CLASSIFICATION tasks**. However, keep in mind that these algorithms can also perform regression!  

This time however, "target" (1 = person has heart disease, 0 = person does not have heart disease) will by our y outcome variable - the others will be our x features. 

```{r}
task_class = list(
  data = data,
  outcome = "target"
)
(task_class$covariates = setdiff(names(task_class$data), task_class$outcome))
```
           
Our factors have still been converted to indicators from the regression setup! :) 

### Stratified random split

We then can use [stratified random sampling](https://stats.stackexchange.com/questions/250273/benefits-of-stratified-vs-random-sampling-for-generating-training-data-in-classi) to divide our data into training and test sets; 70% of the data will be assigned to the training set and the remaining 30% will be assigned to the holdout, or test, set. 

```{r}
# Set seed for reproducibility.
set.seed(2) 

# Create a stratified random split.
training_rows =
  caret::createDataPartition(task_class$data[[task_class$outcome]],
                             p = 0.70, list = FALSE) 

# Partition training dataset
task_class$train_rows = training_rows

mean(task_class$data[training_rows, "target"])
table(task_class$data[training_rows, "target"])

mean(task_class$data[-training_rows, "target"])
table(task_class$data[-training_rows, "target"])
```

### Save our preprocessed data

We save our preprocessed data into an RData file so that we can easily load it in other files.

```{r save_data}
save(data, task_reg, task_class,
     file = "data/preprocessed.RData")
```
