# OLS and lasso

# OLS in 8 steps

Below is an refresher of ordinary least squares linear (OLS) regression that predicts age using the other variables as predictors. 

```{r}
# Look at Y and X variables
head(train_val_reg, n = 20)

head(train_x_reg, n = 20)

# Fit the regression model; lm() will automatically add a temporary intercept column
reg_fit = lm(train_val_reg ~ ., data = as.data.frame(train_x_reg))

# View the output
summary(reg_fit) 

# Predict outcome for the training data
reg_predicted = predict(reg_fit, as.data.frame(test_x_reg))

# 8. Calculate mean-squared error
MSE_reg = mean((test_val_reg - reg_predicted)^2)

MSE_reg

sqrt(MSE_reg) # RMSE
```

# Lasso

Fit the lasso! However, because glmnet expects a matrix of predictors, use `as.matrix` to convert it from a data frame to a matrix. 

```{r}
lasso = cv.glmnet(train_x_reg, train_val_reg, family = "gaussian", alpha = 1)
```

Visualize the distribution of lamba. 
```{r}
plot(lasso)
```

Show plot of different lambda values: 
```{r}
plot(lasso$glmnet.fit, xvar = "lambda", label = T)
```

Show the minimum MSE:
```{r}
lasso$lambda.min
```

Show higher lambda within [one standard error](https://stats.stackexchange.com/questions/80268/empirical-justification-for-the-one-standard-error-rule-when-using-cross-validat) of performance of the minimum
```{r}
lasso$lambda.1se
```

Look at the coefficients
```{r}
coef(lasso, s = "lambda.1se")
```

Look at the coefficients for lambda.min
```{r}
coef(lasso, s = "lambda.min")
```

Predict on the test set
```{r}
predictions = predict(lasso, s = lasso$lambda.1se, newx = as.matrix(test_x_reg))
```

Calculate MSE and RMSE:
```{r}
# Calculate mean-squared error.
mean((predictions - test_val_reg)^2)

# Calculate root mean-squared error.
sqrt(mean((predictions - test_val_reg)^2))
```

**Big question 1:** What are the statistical assupmtions for OLS regression? What are the statistical assumptions for Lasso? 

##### Challenge 1
Construct a lasso to predict the "Petal.Width" variable from the iris dataset. What predictors are most strongly associated with values for "Petal.Width"? 