---
title: "Introduction to Machine Learning in R"
author: "Evan Muzzall and Chris Kennedy"
date: "4/23/2018"
output:
  html_document:
    toc: yes
    toc_float: yes
---

# Introduction 
Be sure to check out the D-Lab [Machine Learning Working Group](http://dlab.berkeley.edu/working-groups/machine-learning-working-group-0) to learn more about machine learning in both R and Python, give a lightning talk, lead a coding walkthrough, and participate in other projects.  

# Resources
James et al (amazon) (free)
- strike car?

# Package installation
```{r, eval = F}
if (FALSE) {
  # Run this line manually to install the necessary packages.
install.packages(c("car", "caret","chemometrics", "ck37r", "class", "devtools", "doParallel", "gbm", "ggplot2", "gmodels", "pROC", "randomForest", "RhpcBLASctl", "rpart", "rpart.plot", "SuperLearner"))
}
  
library(car)
library(caret)
library(chemometrics)
library(ck37r)
library(class)
library(doParallel)
library(gbm)
library(ggplot2)
library(gmodels)
library(parallel)
library(pROC)
library(randomForest)
library(RhpcBLASctl)
library(rpart)
library(rpart.plot)
library(SuperLearner)
```

# Brief history of machine learning
Machine learning evolved from scientific pursuits in computational/information theory, artificial intelligence, and pattern recognition.  

How to define machine learning?  
1) **In general:** algorithms, computers, and other machines that can "learn" without direct input from a programmer.  
2) **Practically:** A suite of tools for investigating/modeling/understanding complex datasets.  
3)  **Specifically:** (see below)

A proto-example:  
- [Pascal's calculator](http://history-computer.com/MechanicalCalculators/Pioneers/Pascal.html)  
Rapid advances:  
- [McCulloch Pitts neuron model](http://wwwold.ece.utep.edu/research/webfuzzy/docs/kk-thesis/kk-thesis-html/node12.html)  
- [Turing test](http://www.jstor.org/stable/pdf/2251299.pdf)  
- [Rosenblatt's perceptron](http://www.ling.upenn.edu/courses/cogs501/Rosenblatt1958.pdf)  
- [Samuels and the game of checkers](http://ucelinks.cdlib.org:8888/sfx_local?sid=google&auinit=AL&aulast=Samuel&atitle=Some+studies+in+machine+learning+using+the+game+of+checkers&id=doi:10.1147/rd.33.0210&title=IBM+Journal+of+Research+and+Development&volume=3&issue=3&date=1959&spage=210&issn=0018-8646)  

Modern topics:  
- [Turing Test: 50 years later](http://www.cs.bilkent.edu.tr/~akman/jour-papers/mam/mam2000.pdf)  
- [computer vision](http://www.sciencedirect.com/science/article/pii/S1071581916301264)  
- [data cleaning](http://www.betterevaluation.org/sites/default/files/data_cleaning.pdf)  
- [machine learning software, tools, and methods](http://achcoop.ca/Data%20Mining%20Practical%20Machine%20Learning%20Tools%20and%20Techniques/)  
- [robotics](https://arxiv.org/abs/1708.04677)  
- [cloud computing](https://arxiv.org/abs/1707.07452)  

The importance of statistics:  
- [Welling's commentary](https://www.ics.uci.edu/~welling/publications/papers/WhyMLneedsStatistics.pdf)  
- [Srivastava's discussion](https://www.analyticsvidhya.com/blog/2015/07/difference-machine-learning-statistical-modeling/)  

Seek "actionable insight":  
- ["actionable insight"](https://www.techopedia.com/definition/31721/actionable-insight)  

# Supervised machine learning
Selecting a machine learning algorithm depends on the characteristics of the problem being investigated. Machine learning is generally divided into three broad classes of learning: [supervised](https://en.wikipedia.org/wiki/Supervised_learning), [unsupervised](https://en.wikipedia.org/wiki/Unsupervised_learning), and [reinforcement](https://en.wikipedia.org/wiki/Reinforcement_learning). In this workshop we will focus on two main subtypes of supervised machine learning: classification and regression.  

The syntax for supervised machine learning algorithms can be thought of like this:  

Y ~ X~1~ + X~2~ + X~3~… X~n~

Y is the dependent/response/target/outcome variable  
X are the independent/input/predictor/feature variables.  

Supervised machine learning methods learn a target function $f$ that best maps X to Y based on a set of training data [training data](https://en.wikipedia.org/wiki/Training,_test,_and_validation_sets). The function _f_ relates X to Y but this is unknown and we must estimate it. Thus, we can predict Y using $\hat{y} = \hat{f}(X)$ for new data (test dataset) and evaluate how well the algorithm learned the target function when introduced to new data. Epsilon $\epsilon$ is the random error, is independent of X, and averages to zero.  

**How to define machine learning? (revisited)**  
More specifically, we can think of machine learning as a bunch of methods to estimate $f$. See [An Introduction to Statistical Learning, Chapters 1 and 2](http://www-bcf.usc.edu/~gareth/ISL/) to learn more.  

##### Classification or regression?
**Classification** is used when the `Y` variable is categorical/discrete. Binary examples generally refer to a yes/no situation where a `1` is prediction of the "yes" category and `0` as the "no". Classification models the probability that the outcome variable is 1 based on the covariates: $Pr(Y = 1 | X)$. This can be extended to multi-level classification as well.  

**Regression** is used when the target Y variable is continuous. Regression models the conditional expectation (conditional mean) of the outcome variable given the covariates: $E(Y | X)$.  

Examples:  
- [text analysis](http://www.digitalhumanities.org/dhq/vol/3/2/000041/000041.html)  
- [network analysis](http://www.digitalhumanities.org/dhq/vol/3/2/000044/000044.html)  
- [real estate](https://arxiv.org/pdf/1609.08399.pdf)  
- [climate science](http://www.geosci-model-dev-discuss.net/gmd-2015-273/gmd-2015-273.pdf)  
- [voter turnout](http://pubs.aeaweb.org/doi/pdfplus/10.1257/jep.28.2.51)  
- [public health](http://journals.plos.org/plosntds/article?id=10.1371/journal.pntd.0004549)  

##### Data preprocessing
A longstanding first step is to split a dataset into **"training"** and **"test"** subsets. A training dataset usually consists of a majority portion of the original dataset so that an algorithm can learn the model. The remaining portion of the dataset is designated to the test dataset to evaluate model performance on data the model has not yet seen.  

> QUESTION: what major issue was not mentioned here? (HINT: see below!)

##### Model performance
**Performance metrics** are used to see how well a model predicts a specified outcome on training and test datasets.  

A model that performs poorly on the training dataset is **underfit** because it is not able to discern relationships between `X` and `Y` variables.  

A model that performs well on the training dataset but poorly on the test dataset is saide to be **overfit** because the model performed poorly when given new data - the patterns found in the test data simply might not exist in the test data.  

### Common performance metrics
*Accuracy  
*Mean squared error  
*Sensitivity and specificity  
*Area under the ROC curve (AUC)  
*Cross validated error  

Learn more:
*[Jason Brownlee - How to Evaluate Machine Learning Algorithms](http://machinelearningmastery.com/how-to-evaluate-machine-learning-algorithms/) and [Assessing and Comparing Classifier Performance with ROC Curves](http://machinelearningmastery.com/assessing-comparing-classifier-performance-roc-curves-2/)  
*[Sensitivity and specificity](https://uberpython.wordpress.com/2012/01/01/precision-recall-sensitivity-and-specificity)
*[An Introduction to Statistical Learning - ROC, pp.147-149](http://www-bcf.usc.edu/~gareth/ISL/ISLR%20Sixth%20Printing.pdf)  
*[R-bloggers - cross validation example](https://www.r-bloggers.com/cross-validation-for-predictive-analytics-using-r/)  
*[Rob J. Hyndman - why every statistician should know about cross-validation](http://robjhyndman.com/hyndsight/crossvalidation/)  

# Workshop goals
### General goals
1) Learn the basics of coding six machine learning algorithms in R:  
  *k-nearest neighbor
  *linear regression
  *decision tree
  *random forest
  *boosting
  *SuperLearner ensemble
2) Examine the performances of these models  
3) Construct ensembles to simultaneously compare multiple algorithms  
4) Vizualize outputs  

### Specific goals
Use the  `PimaIndiansDiabetes2` dataset from the [`mlbench` package](https://cran.r-project.org/web/packages/mlbench/mlbench.pdf) to investigate the following questions:  

1) **Binary classification examples:** How reliably can machine learning algorithms predict a person's diabetes status using the other variables?  
2) **Regression example:** How well can a person's age be predicted using the other variables?  

# Load the data
Load the `PimaIndiansDiabetes2` and `iris` datasets
```{r}
library(mlbench)
# load the Mroz dataset
data("PimaIndiansDiabetes2") 

# read variable descriptions
?PimaIndiansDiabetes2

# rename the dataset to something simpler to type (pidd = "Pima Indians Diabetes Dataset")
pidd = PimaIndiansDiabetes2 

# view basic info
str(pidd) 

# also load iris dataset for challenge questions
data(iris)
str(iris)
?iris
```

##### Data preprocessing
Drata peprocessing is an integral first step in machine learning workflows. Because different algorithms sometimes require the moving parts to be coded in slightly different ways, always make sure you research the algorithm you want to implement so that you properly identify your Y and X variables and split your data into training and test sets when neeeded.  

> One additional preprocessing aspect to consider: datasets that contain factor (categorical) features should be expanded out into numeric indicators (this is often referred to as [one-hot encoding](https://hackernoon.com/what-is-one-hot-encoding-why-and-when-do-you-have-to-use-it-e3c6186d008f). You can do this manually with the `model.matrix` R function. This makes it easier to code a variety of algorithms to a dataset as many algorithms handle factors poorly (decision trees are the main exception). When we predict age in our regression example below, we will do this manually for good practice although functions like the `lm` function we will use will automatically expand the `diabetes` factor predictor into do numeric indicators.  

- **k-nearest neighbor:** we want to remove our Y variable from our dataframe of X features and store it as its own object and split the data into training and test sets. 
- **linear regression:** we will similarly remove our Y outcome from our dataframe of X predictors, but this time will both split the data into training and test sets _and_ we will 1) train the same model on _all of the data_ but then also 2) use R's handy `predict` function to create a test set based on predicted values from all of the data. 

> NOTE: Keep in mind that training/test dataset splitting is not always necessary. We will introduce you to cross-validation in the second half of this workshop. 

##### Handling missing data
First, like many "tasks" involving computing, statistics, and number theory, missing values need to be handled somehow. Listwise deletion (deleting any row with at least one missing case) is common but this method throws out a lot of useful informaiton. Many advocate for mean imputation, but arithmetic means are often too sensitive to outliers. Still, others advocate for Chained Equation/Bayesian/Expectation Maximization imputation (e.g., the [mice](http://www.stefvanbuuren.nl/publications/mice%20in%20r%20-%20draft.pdf) and [Amelia II](https://gking.harvard.edu/amelia) R packages).  

However, median imputation is demonstrated below for the classification examples (we will use listwise deletion for our regression example). 
```{r}
# First, count the number of missing cases in our pidd dataset
sum(is.na(pidd)) # 652 missing cases

# Then, compute the proportion of missing cases in pidd
sum(is.na(pidd)) / (nrow(pidd) * ncol(pidd)) # ~9% of pidd is missing
```

Now, median impute the missing values! We also want to create missingness indicators to inform us about the location of missing data. Thus, we will have some additional columns to our data frame. 

First, 
```{r}
library(ck37r)
outcome = "diabetes"

# Impute missing data and add missingness indicators.
# Don't impute the outcome though.
result = ck37r::impute_missing_values(data, skip_vars = outcome)
# Use the imputed dataframe.
pidd = result$data

# view new columns
str(pidd)

# No more missing data! 
sum(is.na(pidd))
```


# Defining Y outcome vectors and X feature dataframes
### Classification setup
Assign outcome to its own vector for **CLASSIFICATION tasks:** k-nearest neighbor, decision tree, random forest, boosting, and SuperLearner algorithms. However, keep in mind that these algorithms can also perform regression!
```{r}
# View pidd variable names
names(pidd)

# 1) Define Y for classification (has diabetes? "pos" or "neg")
Y = pidd$diabetes
head(Y, n = 20)

# 2) Then, convert "pos" to 1 and "neg" to 0
Y = ifelse(Y == "pos", 1, 0)
head(Y, n = 20)

# 3) Finally, define the X feature/predictor dataframe that excludes Y
data = subset(pidd, select = -diabetes)
head(data) # "diabetes" column has been successfully removed
```

We then can take the "traditional" approach to data splitting and divide our data into training and test sets; 70% of the data will be assigned to the training set and the remaining 30% will be assigned to the holdout, or test, set. 
```{r}
library(caret)
# set seed for reproducibility
set.seed(1)

# Create a stratified random split
classification_split <- createDataPartition(Y, p=0.70, list=FALSE) 
classification_split

train_X <- data[classification_split, ] # partition training dataset
test_X <- data[-classification_split, ] # partition test dataset

train_label <- Y[classification_split] # partition training Y vector labels
test_label <- Y[-classification_split] # partition test Y vector labels
```

### Regression setup
Now we want to do something similar for our **REGRESSION task:** linear regression. There are forms of regression that can perform classification (such as `glm`), but the one we use here - `lm` - is not one of them. 
```{r}
# View pidd variable names
names(pidd)
pidd_reg = na.omit(pidd)
dim(pidd_reg)
dim(pidd)

# 4) Define Y for regression ("age")
Y_reg = pidd$age
head(Y_reg, n = 20)

# 5) Then, define the X feature/predictor dataframe that excludes Y_reg
data_reg = subset(pidd, select = -age)
head(data_reg) # "age" column has been successfully removed - but "diabetes" column is now present! 

# 6) Finally, convert "diabetes" feature into numeric indicators
data_ind = data.frame(model.matrix( ~ . -1, data_reg)) 
head(data_ind) # what happened to the "diabetes" column?
```

Now we want to do the same splitting process like we did for classification above: 70% of data will be assigned to the training set and 30% to the test set. 
```{r}
library(caret)
# set seed for reproducibility
set.seed(1)

# Create a stratified random split - looks very similar to above, but this time is for regression
regression_split <- createDataPartition(Y_reg, p=0.70, list=FALSE) 
regression_split

train_X_reg <- data[regression_split, ] # partition training dataset for regression
train_X_reg <- data[-regression_split, ] # partition test dataset for regression

train_label <- Y[classification_split] # partition training Y vector labels
test_label <- Y[-classification_split] # partition test Y vector labels
```

# Algorithm walkthroughs
#1. k-nearest neighbor (knn)
The k-nearest neighbor (knn) makes no assumptions about the underlying distribution of the data. To classify a data point, knn uses the characteristics of the points around it to determine how to classify it. [Euclidean distances](https://en.wikipedia.org/wiki/Euclidean_distance) between points are used in this example.  

`k` is the number of neighbors used to classify the point in question. Choosing a proper "k" is integral to finding the best-performing model. *Large k-values* could be bad because the class with the largest size might win regardless of influence of more proximate points. *Small k-values* might also be bad because neighboring points might become overly influential.  

#####1.2 knn - selecting a starting "k-value"
Different methods exist for choosing a start point for "k". Let's use the square root of the number of rows in the training datset: 
```{r}
round(sqrt(nrow(train_X)), digits=2) # 15.17
```

Fit the model! Our goal is to predict the classification accuracy of our `Y` variable (yes/no [rather, 1/0] whether or not a person has diabetes based on the other `train_X` predictor variables:
```{r}
library(class)
set.seed(1)
data_predicted <- knn(train = train_X, test = test_X,
                     cl = train_label, k = 15, prob = TRUE)
```

Using accuracy as our performance metric, compute a contingency table to see how well the model predicted yes and no:
```{r}
library(gmodels)
CrossTable(x = test_label, y = data_predicted, 
           prop.chisq = F,
           prop.r = F,
           prop.c = F,
           prop.t = F)

# Compute accuracy
mean(test_label == data_predicted) # ~ 0.52
```
How did it do?  

##### 1.4 knn - improving performance metrics
Two common ways of improving model performance are to 1) standardize the data so that scale does not unduly influence classification, and 2) change the k-value. 

##### 1.4.1 knn - improving performance metrics - scale the data
Scaling the data is useful so that variables with large values do not bias the prediction. Create a new copy of the "data" dataset called "data_scaled", which will contain scaled values with means equal to 0 and standard deviations equal to 1. Our `Y` variable remains unchanged:
```{r}
data_scaled <- scale(data, center = TRUE, scale = TRUE)

# Look at the new distribution.
summary(data_scaled)
```

Repeat the process:  
1) split the scaled data  
```{r}
# Use "split" to create the training partition.
set.seed(1)
train_scaled <- data_scaled[split, ] 

# Create the test partition.
test_scaled <- data_scaled[-split, ] 

# Extract outcome data for training set.
train_label_scaled <- Y[split] 

# Extract outcome data for test set.
test_label_scaled <- Y[-split] 
```

2) fit the model on the scaled data  
```{r}
library(class)
set.seed(1)
data_predicted_scaled <- knn(train = train_scaled,
                            test = test_scaled, 
                            cl = train_label_scaled,
                            k = 23, prob = TRUE)
```

3) examine the accuracy of predictions on the scaled data:
```{r}
library(gmodels)
CrossTable(x = test_label_scaled, y = data_predicted_scaled, 
           prop.chisq = F,
           prop.r = F,
           prop.c = F,
           prop.t = F)

# Compute accuracy
mean(test_label_scaled == data_predicted_scaled) # 0.68
```
What happened?  

##### 1.4.2 knn - improving performance metrics - change k-value
It also helps to investigate the cross-validated errors for multiple k-values at once to see which is ideal. Plot the cross-validated errors:
```{r}
library(chemometrics)
set.seed(1)
knn_k <- knnEval(data_scaled, Y, split, 
                 knnvec=seq(1, 51, by = 1), 
                 legpo="bottomright", las = 3)
```
A combination of scaling the data and searching for the best cross-validated "k" is ideal.  

### Challenge 1  
1. What happens to accuracy when k = 2, 10, or 20?  
2. What happens when you set k to 17, the optimal value from the cross-validation plot? What happens to the predictive accuracy when you switch k to this value?  
3. Using what you learned above, classify predictive accuracies of the `Species` variable in the  "iris" dataset. 

# 2.  Linear regression
- Ordinary least squares regression (OLS) can be used when the target Y variable is continuous.  
- Remember that under the hood, `lm` and `glm` are one-hot encoding factors to indicators. Write it out again for good practice: 
```{r}
# View structure of Mroz.
str(Mroz)

# Define our continuous Y response variable, "lwg". Call it "Y2".
Y2 <- Mroz$lwg

# Expand our predictor X variables out into indicators.
data2 <- data.frame(model.matrix( ~ ., subset(Mroz, select = -lwg)))
names(data2)

# Remove intercept.
data2 <- data2[, -1]

# View structure of "data2".
str(data2)
```

- Fit a model that predicts the `Y` variable log wage gap (`lwg`) using the other variables as predictors.  

- Mean squared error (MSE) will be our performance metric. MSE measures the difference between observed and expected values, with smaller values tending to reflect greater predictive accuracy. 
```{r}
# Fit the regression model.
data2_lm <- lm(Y2 ~ ., data = data2)

# View the regression results.
summary(data2_lm) 

# Predict the outcome back onto the training data.
data2_predicted <- predict(data2_lm, data2) 

# Calculate mean-squared error. 
MSE <- mean((Y2 - data2_predicted)^2)   

MSE
sqrt(MSE)
```
See the [yhat blog for Fitting & Interpreting Linear Models in R](http://blog.yhat.com/posts/r-lm-summary.html) to learn more about `lm` output.  

### Challenge 2
1) What happens to MSE when you try to predict "age" and "lfp"?  
2) Set up a regression model that predicts one of the numeric variables from the "iris" dataset. 

# 3.  Decision trees 
- Decision trees are recursive partitioning methods that divides the predictor spaces into simpler regions and can be visualized in a tree-like structure. Decesion trees attempt to classify data by dividing it into subsets according to a `Y` output variable and based on some predictors.  

- Let's see how a tree-based method classifies women's participation in the workforce.  

Note that we do not have to use `model.matrix` for our decision tree algorithm here:
```{r}
library(rpart)
# Choose method="anova" for a regression tree
# Instead of gini coefficient, select "information" inside the "parms" parameter
dt <- rpart(Mroz$lfp ~ ., data = subset(Mroz, select = -lfp),
            method = "class",
            parms = list(split = 'information'))

# Here is the text-based display of the decision tree. Yikes! 
print(dt)

# The plot is much easier to interpret.
library(rpart.plot)
rpart.plot(dt, cex = 1, type = 4, extra = 101)
           # shadow.col = "lightgray", 
           
# Variable importance is also informative:
dt$variable.importance
```

In decision trees the main hyperparameter (configuration setting) is the *complexity parameter* (CP). A high CP results in a simple decision tree with few splits, whereas a low CP results in a larger decision tree with many splits. So, the name is a little counterintuitive.  

rpart uses cross-validation internally to estimate the accuracy at various complexity parameter (CP) settings. We can review those to see what setting seems best.
```{r}
# Print the results for various CP settings.
# Here we want the setting with the lowest "xerror".
printcp(dt)

# Plot the performance estimates for different CP settings.
# We see that a larger tree (lower CP) seems more accurate.
plotcp(dt)

# We can "prune" a tree to a higher CP setting if we want.
# Note: due to rounding we make the cp value slightly higher than
# was shown in the table.
tree_pruned <- prune(dt, cp = 0.033847)

# Print detailed results, variable importance, and summary of splits.
summary(tree_pruned) 
rpart.plot(tree_pruned)

# Also see the "control" argument inside the rpart function: 
?rpart
```

Here, each node shows the cutoff for each of the predictors and how it contributed to the split. View a helpful decision tree tutorial at [GormAnalysis](https://gormanalysis.com/decision-trees-in-r-using-rpart/).  

### Challenge 3
1. Make a decision tree that tries to classify something other than "lfp".  
2. What are the "minsplit" and "cp" hyperparameters within the "control" parameter? How might you go about figuring out what these are?  
HINT: the syntax might look like this: `dt2_control <- rpart.control(minsplit = 20, cp = 0.001)`  

# 4. Random forests
Random forest is an algorithm that seeks to improve on the performance of a single decision tree by taking the average of many trees. Thus, a random forest is an **ensemble** method, or model averaging approach. The algorithm was invented by Berkeley's own Leo Breiman in 2001, who was also a co-creator of decision trees (see 1984 CART book).

Random forests are an extension of **bagging**, in which multiple samples of the original data are drawn with replacement (aka "bootstrap samples"), an algorithm is fit separately to each sample, then the average of those estimates is used for prediction. While bagging can be used any algorithm, RF uses decision trees as its base learner. Random Forests add another level of randomness by also randomly sampling the features (or covariates) at each split in each decision tree. This makes the decision trees use different covariates and therefore be more unique. As a result, the average of these trees tends to be more accurate overall.

```{r}
# View "Y" and "data"
table(Y)
str(data)

# split this data, similar to before
str(train_X)
str(test_X)
```

Now, let's fit a model that tries to predict the number of women who participated in the labor force in 1975 again using the other variables as our X predictors: 
```{r}
library(randomForest)
set.seed(1)
rf1 <- randomForest(train_label ~ ., 
                    data = train_X, 
                    # Number of trees
                    ntree = 500, 
                    # Number of variables randomly sampled as candidates at each split.
                    mtry = 2, 
                    # We want the importance of predictors to be assessed.
                    importance = TRUE) 

rf1

```
The "OOB estimate of error rate" shows us how accurate our model is. $accuracy = 1 - error rate$.  OOB stands for "out of bag" - and bag is short for "bootstrap aggregation". So OOB estimates performance by comparing the predicted outcome value to the actual value across all trees using only the observations that were not part of the training data for that tree.

We can examine the relative variable importance in table and graph form. Random Forest estimates variable importance by separately examining each variable and estimating how much the model's accuracy drops when that variable's values are randomly shuffled (permuted). The shuffling temporarily removes any relationship between that covariate's value and the outcome. If a variable is important then the model's accuracy will suffer a large drop when it is randomly shuffled. But if the model's accuracy doesn't change it means the variable is not important to the model - e.g. maybe it was never even chosen as a split in any of the decision trees.
```{r}
# As the function name suggests, this creates a variable importance plot.
varImpPlot(rf1)

# Raw data.
rf1$importance
```
You can read up on the [gini coefficient](https://en.wikipedia.org/wiki/Gini_coefficient) if interested. It's basically a measure of diversity or dispersion - a higher gini means the model is classifying better. The gini version does not randomly shuffle the variable cells.  

Now, the goal is to see how the model performs on the test dataset:
```{r}
rf_pred <- predict(rf1, newdata = test_X)
table(rf_pred, test_label)
```

Check the accuracy of the test set:
```{r}
mean(rf_pred == test_label) # 0.7377778
# devtools::install_github("ck37/ck37r")
summary(ck37r::rf_count_terminal_nodes(rf1))
```
How did it do? Are the accuracies for the training and test sets similar?  

### Challenge 4

1. Try a few other values of mtry - can you find one that has improved performance?
2. Maxnodes is another tuning parameter for randomForest - does changing it improve your performance?
3. Select another Y variable to predict and evaluate performance on the train_rf and test_rf. How is the fit of the model for this variable?  

# 5. Boosting

"Boosting is a general method for improving the accuracy of any given learning algorithm" and evolved from AdaBoost and PAC learning (p. 1-2). Gradient boosted machines are ensembles decision tree methods of "weak" trees that are just slightly more accurate than random guessing. These are then "boosted" into "strong" learners. That is, the models don't have to be accurate over the entire feature space.  

The model first tries to predict each value in a dataset - the cases that can be predicted easily are _downweighted_ so that the algorithm does not try as hard to predict them.  

However, the cases that the model has difficulty predicting are _upweighted_ so that the model tries more assertively to predict them. This continues for multiple "boosting iterations", with a training-based performance measure produced at each iteration. This method can drive down generalization error (p. 5). (from [Freund Y, Schapire RE. 1999. A short introduction to boosting. Journal of Japanese Society for Artificial Intelligence 14:771-780](https://cseweb.ucsd.edu/~yfreund/papers/IntroToBoosting.pdf).  

Remember, our data are still split from before:
```{r}
# view "Y" and "data"
table(Y)
str(data)
nrow(train_X) + nrow(test_X) == nrow(data) # sanity check
```

Rather than testing only a single model at a time, it is useful to compare tuning parameters within that single model. Bootstrap is often a default, but we want cross-validation. We also want to specify different tuning parameters.  

Let us first create two objects - `control` and `grid`. `control` will allow us to tune the cross-validated performance metric, while `grid` lets us evaluate the model with different characteristics: 
```{r}
# Choose 10-fold repeated measure cross-validation as our performance metric
# (instead of the default "bootstrap").
control <- trainControl(method = "repeatedcv", 
	repeats = 2,
	# Calculate class probabilities.
	classProbs = TRUE,
	# Indicate that our response varaible is binary.
	summaryFunction = twoClassSummary) 

grid <- expand.grid(
  # Number of trees to fit, aka boosting iterations.
  n.trees = seq(500, 1500, by = 500),
  # Depth of the decision tree.
	interaction.depth = c(1, 3, 5), 
  # Learning rate: lower means the ensemble will adapt more slowly.
	shrinkage = c(0.001, 0.01, 0.1),
  # Stop splitting a tree if we only have this many obs in a tree node.
	n.minobsinnode = 10)
grid
```

Fit the model. Note that we are now using the area under the ROC curve as our performance metric. This can be graphed for quick visualization: 
```{r}
library(caret)
library(pROC)
set.seed(1)
label_numeric = as.numeric(train_label == "yes")
table(label_numeric, train_label)

# cbind: caret expects the Y response and X predictors to be part of the same dataframe
gbm1 <- train(train_label ~ ., data = cbind(train_label, train_X), 
	method = "gbm",
	# Use AUC as our performance metric, which caret incorrectly calls "ROC".
	metric = "ROC",
	trControl = control,
	tuneGrid = grid,
	# Keep output more concise.
	verbose = T)

# See how long this algorithm took to complete.
gbm1$times 

# Review model summary table.
gbm1 

# Plot variable importance.
summary(gbm1, las = 2)

# Generate predicted values.
gbm.pred <- predict(gbm1, test_X)

# Generate class probabilities.
gbm.prob <- predict(gbm1, test_X, type = "prob")

# Define ROC characteristics.
rocCurve <- roc(response = test_label,
	predictor = gbm.prob[, "yes"],
	levels = rev(levels(test_label)),
	auc=TRUE, ci=TRUE)

# Plot ROC.
plot(rocCurve, print.thres = "best", main = "GBM", col = "blue") 
# ggsave("gbm AUC.png")

# Plot the cross-validated AUC of the different configurations.
ggplot(gbm1) + theme_bw() + ggtitle("GBM model comparisons") 
# ggsave("gbm tuning comparison.png")
```

Also check out the ["xgboost" R package](https://cran.r-project.org/web/packages/xgboost/index.html) for extreme gradient boosting.  

### Challenge 5

1. Using your Mroz dataset, fit a GBM model where "wc" is your Y variable. What happens to variable relative importance? What happens to AUC?

# 6. Ensemble methods

The ["SuperLearner" R package](https://cran.r-project.org/web/packages/SuperLearner/index.html) is a method that simplifies ensemble learning by allowing you to simultaneously evaluate the cross-validated performance of multiple algorithms and/or a single algorithm with differently tuned hyperparameters.  

Let's see how the five algorithms you learned in this workshop (kNN, linear regression, decision trees, random forest, and gradient boosted machines) compare to each other and to the mean of Y as a benchmark algorithm, in terms of their cross-validated error!

A "wrapper" is a short function that adapts an algorithm for the SuperLearner package. Check out the different algorithm wrappers offered by SuperLearner:
```{r}
library(SuperLearner)
listWrappers()
```

Instead of splitting the data like before, since we are using cross-validation we actually want to use the entire `Mroz` dataset:
```{r}
str(Mroz)

# convert our Y variable to numeric type
Mroz$lfp <- ifelse(Mroz$lfp == "yes", 1, 0)

# Now, tell SuperLearner which algorithms to urn. In the interest of time we are not including GBM, but "SL.gbm" is the wrapper.
SL_library <- c("SL.mean", "SL.knn", "SL.glm", "SL.rpart", "SL.randomForest")

```

Fit the ensemble:

```{r}
library(SuperLearner)
# This is a seed that is compatible with multicore parallel processing.
# See ?set.seed for more information.
set.seed(1, "L'Ecuyer-CMRG") 

# This may take a minute to execute.
cv_sl <- CV.SuperLearner(Y = Mroz$lfp, X = data,
                         SL.library = SL_library,
                         family = binomial(),
                         # Fit only 5 folds for demo purposes only.
                         # For a real analysis we would do 20 folds.
                         cvControl = list(V = 5),
                         # Set to T to show details.
                         verbose = F)
summary(cv_sl)
```

Risk is a performance estimate - it's the average loss, and loss is how far off the prediction was for an individual observation. The lower the risk, the fewer errors the model makes in its prediction. SuperLearner's default loss metric is squared error $(y_{actual} - y_{predicted})^2$, so the risk is the mean-squared error (just like in ordinary least _squares_ regression). View and plot results:
```{r}
summary(cv_sl)

plot(cv_sl) + theme_bw()

# ggsave("SuperLearner.pdf")
```
"Discrete SL" is when the SuperLearner chooses the single algorithm with the lowest risk. "SuperLearner" is a weighted average of multiple algorithms, or an "ensemble". In theory the weighted-average should have a little better performance, although they often tie. In this case we only have a few algorithms so the difference is minor.

### Challenge 6. 
1. If you set the seed again and re-run `CV.SuperLearner` do you get exactly the same results?
2. Choose your favorite 2 algorithms and re-run the SuperLearner with just those algorithms in the library. Does the performance change?
3. What are the elements of the CV_SL object? Take a look at 1 or 2 of them. Hint: use the `names()` function to list the elements of an object, then `$` to access them (just like a dataframe).

A longer tutorial on SuperLearner is available here: https://github.com/ck37/superlearner-guide