---
title: "Participant-instructions"
author: "D-Lab"
date: "3/23/2019"
output: html_document
---

# Participant instructions - Machine Learning in R

Please install the following packages before the workshop date. Run the below code. This code can also be found in the 01-overview.Rmd file in this repository.

```{r}
install.packages(c(# Algorithms
                 "gbm", "glmnet", "randomForest", "ranger", "rpart", "xgboost",
                 # Visualization
                 "ggplot2", "rpart.plot", 
                 # Machine learning frameworks
                 "caret", "SuperLearner",
                 # R utility packages
                 "remotes", "dplyr",
                 # Misc
                 "mlbench", "pROC"))
```

If you are prompted by the question, "Do you want to install from sources the package which needs compilation? (Yes/no/cancel)", type no in your console and press Enter.

... and, the most recent version of ck37r. Run this line of code please! 

```{r}
remotes::install_github(c("ck37/ck37r"))
```

Then, open the slides.html file in the docs folder to get started - contact us with any questions!