# Challenge 1 - solution
# 03-lasso.Rmd 

Fill in the blanks to complete the below lasso code that predicts "Petal.Width" variable from the iris dataset. What predictors are most strongly associated with values for "Petal.Width"? 

### 1. Setup
- Load the iris dataset 
- Load the glmnet library
- Set your seed to 1 for reproducibility

```{r}
data(iris)
library(glmnet)
set.seed(1)
```

### 2. Define y variable

```{r}
iris_y = iris$Petal.Width
```

### 3. Define x variables

Remove Petal.Width from the dataset

```{r}
iris_x = subset(iris, select = -Petal.Width)
dim(iris_x)
```

### 4. Convert Species from factor to indicator

```{r}
?model.matrix
features_iris = data.frame(model.matrix( ~ . -1 , iris_x))
head(features_iris)
```

### 5. Perform a 70/30 random split.

```{r}
(training_size = floor(0.70 * nrow(features_iris)))
# Set seed for reproducibility.
set.seed(1)
training_rows_iris = sample(nrow(features_iris), size = training_size)
```

### 6. Partition training and test sets

```{r}
train_x_iris = features_iris[training_rows_iris, ] # partition training dataset
test_x_iris = features_iris[-training_rows_iris, ] # partition test dataset

train_y_iris = iris_y[training_rows_iris] # partition training Y vector labels
test_y_iris = iris_y[-training_rows_iris] # partition test Y vector labels

### Check lengths of x and y data
dim(train_x_iris)
length(train_y_iris)

dim(test_x_iris)
length(test_y_iris)
```

### 7. Fit lasso model and visualize lambda distribution and different lambda values

```{r}
lasso_iris = cv.glmnet(as.matrix(train_x_iris), train_y_iris, family = "gaussian", alpha = 1)
plot(lasso_iris)
plot(lasso_iris$glmnet.fit, xvar = "lambda", label = T)

# View minimum MSE and higher lambda within one standard error
lasso_iris$lambda.min
lasso_iris$lambda.1se

# View coefficients
coef(lasso_iris, s = "lambda.min")
coef(lasso_iris, s = "lambda.1se")

# Predict on test set
predictions_iris = predict(lasso_iris, s = lasso_iris$lambda.1se, newx = as.matrix(test_x_iris))
```

# 8. Calculate MSE and RMSE

```{r}
# Calculate mean-squared error = 0.0262426
mean((predictions_iris - test_y_iris)^2)

# Calculate root mean-squared error = 0.1619957
sqrt(mean((predictions_iris - test_y_iris)^2))
```
