# Challenge 2
# 04-decision-trees.Rmd

Below is the decision tree code in summarized form from notebook 04-decision-tree.Rmd that classifies if a person suffers from heart disease or not. 

Use the help file for `rpart.control` to figure out what the "minsplit", "cp", and "minbucket" hyperparameters do. Can you improve the cross-validated error by adjusting the configuration of `rpart.control`? 

> NOTE: this might change the number of splits in your tree from 04-decision-trees.Rmd

### 1. Setup

```{r load_packages}
library(ggplot2)
library(rpart)
library(rpart.plot)
```

### 2. Load data from 02-preprocessing.Rmd

```{r setup_data}
# Objects: task_reg, task_class
load("data/preprocessed.RData")
```

### 3. Fit the tree

```{r}
set.seed(3)
tree = rpart::rpart(train_y_class ~ ., data = train_x_class,
             # Use method = "anova" for a continuous outcome.
             method = "class",
             # Can use "gini" for gini coefficient.
             parms = list(split = "information"))
               
# View confusing tree output
print(tree)
```

### 4. Plot the tree and view variable importance

```{r plot_tree}
rpart.plot::rpart.plot(tree) 
names(tree)
tree$variable.importance
```

### 5. Show and plot estimated error rate at different complexity parameter settings.

```{r plotcp_tree}
# Show estimated error rate at different complexity parameter settings.
printcp(tree)

# Plot those estimated error rates.
plotcp(tree)
```

### 6. Prune the tree with `rpart.control`

Select the CP with the lowest xerror (or a simpler tree within one standard error), add it to `rpart.control`, and then refit the tree. 

```{r}
ctrl = rpart.control(minsplit = _, minbucket = _, cp = ) 
```

```{r}
set.seed(3)
tree2 = rpart::rpart(train_y_class ~ ., data = train_x_class,
             # Use method = "anova" for a continuous outcome.
             method = "class",
             # Can use "gini" for gini coefficient.
             parms = list(split = "information"), 
             # Add rpart.control
             _______ = ____)
               
print(tree2)
```


### 7. Print detailed results, variable importance, and summary of splits.

```{r}
summary(tree2) 
rpart.plot(tree2)
```

