# Challenge 1 
# 03-lasso.Rmd 

Fill in the blanks to complete the below lasso code that predicts "Petal.Width" variable from the iris dataset. What predictors are most strongly associated with values for "Petal.Width"? 

### 1. Setup
- Load the iris dataset 
- Load the glmnet library
- Set your seed to 1 for reproducibility

```{r}
data(____)
library(______)
___.____(_)
```

### 2. Define y variable
```{r}
iris_y = iris$_____._____
```

### 3. Define x variables
Remove Petal.Width from the dataset
```{r}
iris_x = subset(iris, ______ = -_____._____)
dim(iris_x)
```

### 4. Convert Species from factor to indicator
```{r}
features_iris = data.frame(_____.______( ~ . -1 , ______))
head(features_iris)
```

### 5. Perform a stratified random split.
```{r}
training_rows_iris = _____::___________________(______, p = 0.70, list = FALSE) 
```

### 6. Partition training and test sets
```{r}
train_x_iris = _____________[__________________, ] # partition training dataset
test_x_iris = _____________[-__________________, ] # partition test dataset

train_y_iris = ______[__________________] # partition training Y vector labels
test_y_iris = ______[-__________________] # partition test Y vector labels

### Check lengths of x and y data
dim(train_x_iris)
length(train_y_iris)

dim(test_x_iris)
length(test_y_iris)
```

### 7. Fit lasso model and visualize lambda distribution and different lambda values
```{r}
lasso_iris = cv.glmnet(as.matrix(____________), train___iris, family = "gaussian", alpha = 1)
plot(lasso_iris)
plot(lasso_iris$glmnet.___, xvar = "lambda", label = T)

# View minimum MSE and higher lambda within one standard error
lasso_iris$lambda.___
lasso_iris$lambda.___

# View coefficients
coef(lasso_iris, s = "lambda.___")
coef(lasso_iris, s = "lambda.___")

# Predict on test set
predictions_iris = predict(lasso_iris, s = lasso_iris$lambda.___, newx = as.matrix(___________))
```

# 8. Calculate MSE and RMSE
```{r}
# Calculate mean-squared error = 0.0262426
mean((predictions_iris - test_y_iris)^2)

# Calculate root mean-squared error = 0.1619957
sqrt(mean((predictions_iris - test_y_iris)^2))
```
